---
title: "Modular Analysis of AS Timeseries Data"
output: html_notebook
  ---

This Notebook is performing modular analysis on The AS time course experiment comparing Immulogical responses from Malaria infection via MT and SBP.

The modules are derived from Singhania et al. (Unpublished) and enrichment is derived using Qusage. 

Qusage will assess the mean expression of all genes in a geneset (module) within a treatment group (MT/SBP) and compare against a control group (Naive), and output the Fold Enrichment of the treatment against the control



Loading Libraries and Data for the Analysis. 

Setting directories for Data
```{r}
library(ComplexHeatmap)
library(tidyverse)
library(qusage)


work.dir    <- '~/projects/AS_infection_timecourse.rnaseq/';
module.dir <- paste(work.dir, "modular_analysis/",sep='');
nf.dir      <- paste(work.dir,"BABS-RNASeq/",sep='');
tmp.dir     <- paste(work.dir,"tmp/",sep='');
data.dir    <- paste(work.dir,"data/",sep='');
results.dir <- paste(work.dir,"results/",sep='');
figures.dir <- paste(results.dir, "figures/",sep='');

module.anno <- read.csv(file=paste(module.dir, 'Annotation of Blood modules.csv', sep=''),
                        header=TRUE,
                        sep=',');
module.dat <- read.csv(file=paste(module.dir, 'Blood modules.csv', sep=''),
                       header=TRUE,
                       sep=',');

rlog.dat <- read.table(file= paste(results.dir, 'genes.rlog_counts.xls', sep=''),
                       header=TRUE,
                       sep='\t');

#List of Module IDs
module.num <- module.anno$X
module.anno
```


Transforming data to create Modules into Geneset format for Qusage

```{r}
module.geneset <- lapply(module.num, function(mod){
  dat <- module.dat[module.dat$Module == mod, 'X']; 
  dat <- dat[!is.na(dat)]
  mod <- dat
})

names(module.geneset) <- module.num


#Check to see if Merge was correct
for (mod in names(module.geneset)){
  anno <- as.numeric(module.anno[module.anno$X == mod, 'Number.of.genes']);
  if (length(module.geneset[[mod]]) == anno){
    print(paste("Geneset", mod, 'contains correct number of genes:', anno, sep=' '))
  } else {
    print('Module gene number does not match')
  }
}

summary(module.geneset)
```

Now we format the rlog data to the format we need

```{r}
#Get Columns we need 
names(rlog.dat);
acute.cols <- grep(pattern = 'd3\\.|d6\\.|d9\\.', x = names(rlog.dat), value = TRUE)
rlog.db <- rlog.dat[ , acute.cols];
rownames(rlog.db) <- rlog.dat$gene_id;

```

Now we set up the first set of Analysis, comparing MT to Naive

```{r}
mvn.cols <- grepl(pattern = 'naive|mosquito', x= names(rlog.db));
mvn.rlog <- rlog.db[, mvn.cols];

results.list <- list()
for (d in c(3,6,9)){
  cols <- grepl(pattern = paste('mosquito\\.d', d, '\\.|naive', sep='')  , x = names(mvn.rlog));
  dat <- mvn.rlog[, cols]
  print(names(dat))
  labels <- sapply(names(dat), function(x){
    ifelse(grepl(pattern = 'mosquito', x = x), 'mt','naive')
  })
  contrast = 'mt-naive';
  results.list[[paste('mvn.d', d, sep='')]] <- qusage(dat, labels, contrast, module.geneset, filter.genes = TRUE);
}
```

Now repeat the same for Blood vs Naive 

```{r}
bvn.cols <- grepl(pattern = 'naive|blood', x= names(rlog.db));
bvn.rlog <- rlog.db[, bvn.cols];

for (d in c(3,6,9)){
  cols <- grepl(pattern = paste('blood\\.d', d, '\\.|naive', sep='')  , x = names(bvn.rlog));
  dat <- bvn.rlog[, cols]
  print(names(dat))
  labels <- sapply(names(dat), function(x){
    ifelse(grepl(pattern = 'blood', x = x), 'sbp','naive')
  })
  contrast = 'sbp-naive';
  results.list[[paste('bvn.d', d, sep='')]] <- qusage(dat, labels, contrast, module.geneset, filter.genes = TRUE);
}
names(results.list)
```

Using combined PDF for testing -- Don't need this anymore 
```{r}
mvn_all = grepl(pattern = 'mvn', x = names(results.list));
mvn_all = results.list[mvn_all]

bvn_all = grepl(pattern = 'bvn', x = names(results.list));
bvn_all = results.list[bvn_all]

# mvn_all[['mvn_combined']] = combinePDFs(mvn_all)
# bvn_all[['bvn_combined']] = combinePDFs(bvn_all)

```


Looking at the overall enrichment of Modules throughout the acute phase

```{r}
# plotCombinedPDF(mvn_all[["mvn_combined"]], path.index=15)
# legend("topleft", legend=c("mvn.d3", "mvn.d6", "mvn.d9", "Meta-Analysis"),
#        lty=1, col=c("red", "blue", "green", "black"))
```



Creating Heatmaps for MT time points 

Create a matrix for all Qusage tables

```{r}
mvn.dat <- list()
n <- 1
for (m in mvn_all){
  name <- names(mvn_all)[n];
  df <- qsTable(m, number = 41)
  df <- df[, c("pathway.name", "log.fold.change")]
  colnames(df) <- c('Module', paste(name, 'logFC', sep='_'))
  mvn.dat[[name]]  <- as.data.frame(df)
  n <- n + 1 
}
mvn.heat <- Reduce(function(x,y) merge(x = x, y = y, by = "Module", no.dups=T), 
       mvn.dat)

# mvn.heat$Module <- module.anno$Biological.process[match(mvn.heat$Module, module.anno$X)]


mvn.heat$anno <- paste(mvn.heat$Module, module.anno$Biological.process[match(mvn.heat$Module, module.anno$X)])
# mvn.heat$anno <- module.anno$Biological.process[match(mvn.heat$Module, module.anno$X)]

mvn.heat

rownames(mvn.heat) <- mvn.heat$Module
mvn.heat <- mvn.heat[,-1]
mvn.heat <- mvn.heat[module.num, ]
rownames(mvn.heat) <- mvn.heat$anno
colnames(mvn.heat) <- c('Day 3', 'Day 6', 'Day 9', 'anno')
# mvn.heat <- mvn.heat[,-4] # Not using for the bubble plot
mvn.heat.fc <- 2^mvn.heat[,1:3]
mvn.heat.fc$anno <- mvn.heat$anno
```

Creating Heatmaps for MvN
```{r}
# # acutemap sample annotation
#     ha1 = HeatmapAnnotation(
#       df  = data.frame(
#         day = as.factor(colnames(mvn.heat))),
#       col =list(
#         day = c(
#           "Day 3" = "#FFFFCC",
#           "Day 6" = "#ADDD8E",
#           "Day 9" = "#31A354",
#           "Acute" = "#006837")
#       )
#     )
# 
# 
# ht <- Heatmap(as.matrix(mvn.heat),
#                   name = "Log Fold Enrichment",
#                   show_row_names          = TRUE,
#                   show_column_names       = TRUE,
#                   column_names_side       = "bottom",
#                   show_column_dend        = FALSE,
#                   cluster_columns         = FALSE,
#                   cluster_rows            = FALSE,
#                   rect_gp                 = gpar(col = "white", lwd = 1),
#                   row_dend_width          = unit(10, "mm"),
#                   column_dend_height      = unit(30, "mm"),
#                   show_heatmap_legend     = TRUE,
#                   column_names_max_height = unit(8, "cm"),
#                   row_names_gp            = gpar(fontsize = 6),
#                   column_names_gp         = gpar(fontsize = 10),
#                   col                     = colorRamp2(c(-1, 0, 1), rev(colorRampPalette(brewer.pal(9, "RdBu"))(3))));
# 
# pdf(paste(results.dir,"/figures/Mosquito_modular_analysis.png",sep=''),width=18,height=25);
# print(ht);
# dev.off();
```



For Blood

```{r}
# plotCombinedPDF(bvn_combined, path.index=14)
# legend("topleft", legend=c("bvn.d3", "bvn.d6", "bvn.d9", "Meta-Analysis"),
#        lty=1, col=c("red", "blue", "green", "black"))
```



Creating Heatmaps for MT time points 

Create a matrix for all Qusage tables

```{r}
bvn.dat <- list()
n <- 1
for (m in bvn_all){
  name <- names(bvn_all)[n];
  df <- qsTable(m, number = 41)
  df <- df[, c("pathway.name", "log.fold.change")]
  colnames(df) <- c('Module', paste(name, 'logFC', sep='_'))
  bvn.dat[[name]]  <- as.data.frame(df)
  n <- n + 1 
}
bvn.heat <- Reduce(function(x,y) merge(x = x, y = y, by = "Module", no.dups=T), 
       bvn.dat)

# bvn.heat$Module <- module.anno$Biological.process[match(bvn.heat$Module, module.anno$X)]

bvn.heat$anno <- paste(bvn.heat$Module, module.anno$Biological.process[match(bvn.heat$Module, module.anno$X)])

rownames(bvn.heat) <- bvn.heat$Module
bvn.heat <- bvn.heat[,-1]
bvn.heat <- bvn.heat[module.num, ]
rownames(bvn.heat) <- bvn.heat$anno
colnames(bvn.heat) <- c('Day 3', 'Day 6', 'Day 9', 'anno')
# bvn.heat <- as.matrix(bvn.heat[, -5])
bvn.heat.fc <- 2^bvn.heat[,1:3]
bvn.heat.fc$anno <- bvn.heat$anno
```


Creating Heatmaps for bvn
```{r}
# # acutemap sample annotation
#     ha1 = HeatmapAnnotation(
#       df  = data.frame(
#         day = as.factor(colnames(bvn.heat))),
#       col =list(
#         day = c(
#           "Day 3" = "#FFFFCC",
#           "Day 6" = "#ADDD8E",
#           "Day 9" = "#31A354",
#           "Acute" = "#006837")
#       )
#     )
# 
# 
# ht <- Heatmap(bvn.heat,
#                   name = "Log Fold Enrichment",
#                   show_row_names          = TRUE,
#                   show_column_names       = TRUE,
#                   column_names_side       = "bottom",
#                   show_column_dend        = FALSE,
#                   cluster_columns         = FALSE,
#                   cluster_rows            = FALSE,
#                   rect_gp                 = gpar(col = "white", lwd = 1),
#                   row_dend_width          = unit(10, "mm"),
#                   column_dend_height      = unit(30, "mm"),
#                   show_heatmap_legend     = TRUE,
#                   column_names_max_height = unit(8, "cm"),
#                   row_names_gp            = gpar(fontsize = 6),
#                   column_names_gp         = gpar(fontsize = 10),
#                   col                     = colorRamp2(c(-1, 0, 1), rev(colorRampPalette(brewer.pal(9, "RdBu"))(3))));
# 
# pdf(paste(results.dir,"/figures/Blood_modular_analysis.png",sep=''),width=18,height=25);
# print(ht);
# dev.off();
# Heatmap(bvn.heat)
# 
# mvn.db

```

```{r}
library(ggplot2)
library(reshape2)
library(RColorBrewer);

# d <- ggplot(diamonds, aes(x = cut, y = clarity))
# 
# d + geom_count(aes(size = stat(prop), group = carat, colour = carat)) +
#   scale_colour_gradient2() +
#   scale_size_area(max_size = 10) 
  

mvn.db <- mvn.heat.fc[ , !names(mvn.heat.fc) %in% 'Acute'];
mvn.db$anno <- factor(mvn.db$anno, levels = mvn.db$anno)
mvn.db <- melt(mvn.db, id.vars = 'anno')

mvn.db$prop <- apply(mvn.db, 1, function(prop){
    mod <- unlist(strsplit(as.character(prop[1]), split= ' '))[1]
    day <- unlist(strsplit(as.character(prop[2]), split= ' '))[2];
    geneset <- module.geneset[[mod]]
    num.genes <- length(geneset)
    res.name <- grep(pattern = paste('mosquito\\.', day, '_vs_naive', sep='')  , x = names(res.list), value = TRUE);
    res <- res.list[[res.name]]
    res <- res[geneset, ]
    res <- res[res$DEG == T,]
    prop <- nrow(res) / num.genes

})
mvn.db
mvn.db[order(mvn.db$prop), ]

x <- ggplot(mvn.db, aes(x = variable, y = reorder(anno, desc(anno)))) + ggtitle('Mosquito Transmission') + xlab('Day') + ylab('Module')

x <- x + geom_count(aes(size = prop, group = value , colour = value), alpha=0.75) +
  labs(size = "Proportion of DEG in Module", colour = 'Enrichment score') +
  scale_color_gradientn(colours = rev(brewer.pal(11, "RdBu")), limits = c(0,2.5)) +
  scale_size_area(max_size = 9) +
  # coord_fixed(ratio=0.25) + 
  theme_bw() +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = 'white'),
    # panel.border = element_rect(fill = "transparent"),
    plot.background = element_blank(),
    legend.background = element_rect(fill = 'transparent'))

ggsave(filename =paste(results.dir,"/figures/mosquito_modular_bubble.png",sep=''), height = 10, width = 8, bg = 'transparent')
```
```{r}
bvn.db <- bvn.heat.fc[ , !names(bvn.heat.fc) %in% 'Acute'];
bvn.db$anno <- factor(bvn.db$anno, levels = bvn.db$anno)
bvn.db <- melt(bvn.db, id.vars = 'anno')

bvn.db$prop <- apply(bvn.db, 1, function(prop){
    mod <- unlist(strsplit(as.character(prop[1]), split= ' '))[1]
    day <- unlist(strsplit(as.character(prop[2]), split= ' '))[2];
    geneset <- module.geneset[[mod]]
    num.genes <- length(geneset)
    res.name <- grep(pattern = paste('blood\\.', day, '_vs_naive', sep='')  , x = names(res.list), value = TRUE);
    res <- res.list[[res.name]]
    res <- res[geneset, ]
    res <- res[res$DEG == T,]
    prop <- nrow(res) / num.genes

})
bvn.db
bvn.db[order(bvn.db$prop), ]

x.blood <- ggplot(bvn.db, aes(x = variable, y = reorder(anno, desc(anno)))) + ggtitle('SBP Infection') + xlab('Day') + ylab('Module')

x.blood <- x.blood + geom_count(aes(size = prop, group = value , colour = value), alpha=0.75) +
  labs(size = "Proportion of DEG in Module", colour = 'Enrichment Score') +
  scale_color_gradientn(colours = rev(brewer.pal(11, "RdBu")), limits = c(0,2.5)) +
  scale_size_area(max_size = 9) + 
  # coord_fixed(ratio=0.4) + 
  theme_bw() +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = 'white'),
    # panel.border = element_rect(fill = "transparent"),
    plot.background = element_blank(),
    legend.background = element_rect(fill = 'transparent'))
# ggsave(filename =paste(results.dir,"/figures/blood_modular_bubble.png",sep=''), height = 10, width = 8)

?element_rect
x
x.blood
x.plot <- ggpubr::ggarrange(x, x.blood, nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
ggsave(filename =paste(results.dir,"/figures/combo_modular_bubble_new.tiff",sep=''), height = 12, width = 14, bg = 'transparent', dpi = 300, units = 'in', device ='tiff', limitsize = TRUE)
?ggsave
x.plot
rm(list = c('x', 'x.plot','x.blood'))
```


