---
title: "EKD11 Analysis"
author: "Chris Alder"
date: "08/11/2019"
output:
  html_document:
      toc: true
      number_sections: true
      toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The Aim of the EKD11 timecourse experiment is to investigate the host's immune response to a primary _P.chabaudi AS_ infection transmitted through different routes; Natural transmission (Mosquito), serially blood passaged (SBP), and recently transmissited blood passage (RTMT). The timecourse looks at the acute phase of the infection, using transcriptomics data taken from days 1, 2, 3, 4 and 6.


# Pre-requisites

Loading the required R library packages and setting the directories for the project analysis
```{r workspace, include=FALSE}
source('/Users/alderc/1-projects/1-PIR_project/3-EKD11_AS_Early_TP/1-R/1-R_scripts/packages.R')

## Directories
nf.dir      <- "/Users/alderc/1-projects/CAMP/1-AS_timecourse/2-EKD11/1-Pipeline/1-Nextflow/";
work.dir    <- "/Users/alderc/1-projects/1-PIR_project/3-EKD11_AS_Early_TP/";
r.dir       <- paste(work.dir, "1-R/1-R_scripts/",sep='');
tmp.dir     <- paste(work.dir,"tmp/",sep='');
data.dir    <- paste(work.dir,"4-data/",sep='');
results.dir <- paste(work.dir,"2-results/",sep='')

for (dir in grep("\\.dir$",ls(),value=T)) {
  if (!file.exists(get(dir))) { dir.create(get(dir),recursive=TRUE, mode="0755"); }
}
```
# Experimental setup

## Loading Genome data

Importing GTF file containing genome information for Mus musculus. GRCm38 version 86

```{r gtf, include=FALSE}
genome.dir <- "~/1-projects/9-Data/1-Reference_genomes/1-Mus_musculus/";
genome     <- "GRCm38";
gtf.file   <- paste(genome.dir,"Mus_musculus.GRCm38.86.gtf",sep='');
gtf.meta    <- import(gtf.file);
gtf.dat    <- as.data.frame(gtf.meta[gtf.meta$type %in% "gene",]);
gtf.dat$GRCm38 <- paste("chr",gtf.dat$seqnames,":",gtf.dat$start,"-",gtf.dat$end,sep='');
rownames(gtf.dat) <- gtf.dat$gene_id;
```

## Design file

This is a file containing all the meta data we need for the data analysis. The important factors for the analysis are: sample, day, delivery, replicate and grp (group)


```{r design, echo=FALSE}
design.file <- paste(nf.dir,"design.csv",sep='');
design      <- read.delim(design.file,header=TRUE,sep=",", stringsAsFactors = TRUE);
rownames(design) <- design$label;
#sets naive to be first level of delivery
design$delivery  <- relevel(design$delivery,"naive");
#orders days in numerical order
design$day       <- factor(as.character(design$day),levels=as.character(sort(unique(design$day))));
design$replicate <- as.factor(design$replicate)

# create a copy of column grp2 and removes .day from naive set so relevel will order all naive as first
design$grp2      <- factor(sub("naive.*","naive",as.character(design$grp)));
design$grp2      <- relevel(design$grp2,"naive");

paged_table(design[ ,c(1,7:11)] , options = list(rows.print = 20))

```

## Importing count data and normalisation

This section will import the results of the RNA-seq pipeline and normalise the data according to library size. A count, log(counts) and TPM data frame will then we created from the normalised results.

Below is the TPM database from the experiment, calculated from RSEM.  
```{r counts, echo = FALSE}
txi <- tximport(paste(nf.dir,"results/rsem/",design$lims.name,".genes.results",sep=''), type="rsem")
txi$length[txi$length==0] <- 1

## Create DESeq2 object
dat <- DESeqDataSetFromTximport(txi, design, ~ grp) # To make Full rank model - Testing with controls removed to check for changes
rowData(dat) <- gtf.dat[rownames(dat),c("gene_id","gene_name","gene_source","gene_biotype","GRCm38")];

## Normalise / scale and stabilise variance
dds <- DESeq(dat);
vsd <- vst(dat, blind = TRUE);


#TPM function
tpm.calc <- function(counts) {
  gene.lengths <- gene.len[rownames(counts), 'length'] /1000
  x <- counts/gene.lengths
  return(t(t(x)*1e6/colSums(x)))
}

# Count, rlog and TPM data frames
count.norm  <- counts(dds,normalized=TRUE);
counts.df <- as.data.frame(count.norm)
vsd.db <- assay(vsd);
tpm.db <- as.data.frame(txi$abundance)
names(tpm.db) <- names(counts.df)

paged_table(tpm.db, options = list(rows.print = 10))
```


#PCA Plot 

PCA plot showing the separation of all samples within the study. The PCA is produced using top 1000 variable genes within the dataset.

```{r pca-3d}

rv <- rowVars(assay(vsd))
select <- order(rv, decreasing = TRUE)[1:1000]

pca <- prcomp(t(assay(vsd)[select, ]))
pca.sum <- summary(pca)
pca.val <- pca.sum$importance[2,1:3]
pca.df <- as.data.frame(pca$x[,1:3])
pca.df <- cbind(pca.df, design[rownames(pca.df), c("delivery", "day", "grp", "label")])
pca.cols <- c("mosquito" = "firebrick", "blood" = "darkblue", "naive" = "#E69F00", 'recent_blood' = 'forestgreen')

p <- plot_ly(data = pca.df) %>%
  add_trace(type = "scatter3d",
            x = ~PC1, y = ~PC2, z = ~PC3,
            color = ~delivery, colors = pca.cols,
            symbols = c("1" = "circle", "2" = "cross", "3" = "x", "4" = "square", "6" = "diamond"),
            text = ~label,
            mode = "markers",
            marker = list(size = 3)) %>% 
  add_trace(type = "scatter3d",
            x = ~PC1, y = ~PC2, z = ~PC3,
            symbol = ~day,
            mode = "markers",
            text = ~label,
            marker = list(color = "grey")) %>% 
  add_trace(type = "scatter3d",
            x = ~PC1, y = ~PC2, z = ~PC3,
            color = ~delivery,
            symbol = ~day,
            mode = "markers",
            text = ~label,
            showlegend = F) %>% 
            layout(scene = list(xaxis = list(title = paste("PC1:", round(pca.val[[1]], 2)*100, "%", sep = " ")),
                                yaxis = list(title = paste("PC2:", round(pca.val[[2]], 2)*100, "%", sep = " ")),
                                zaxis = list(title = paste("PC2:", round(pca.val[[3]], 2)*100, "%", sep = " "))))

# p
```

I've made the PCA in a 3D scatter plot to try and see if there is any separation of the samples before day 3 from PC3 that doesn't seem to appear when using just PC1 and PC2. PC3 only accounts for ~14% of the variance in the dataset, so I wasn't expecting a large change, nonetheless I think the separation it does show is worth keeping. I've also added 2D plots of PC1 vs PC2, PC1 vs PC3 and PC2 vs PC3, which may be more useful. 

```{r pca-1}
p.1 <- plot_ly(data = pca.df) %>%
  add_trace(type = "scatter",
            x = ~PC1, y = ~PC2,
            color = ~delivery, colors = pca.cols,
            symbols = c("1" = "circle", "2" = "cross", "3" = "x", "4" = "square", "6" = "diamond"),
            text = ~label,
            mode = "markers") %>% 
  add_trace(type = "scatter",
            x = ~PC1, y = ~PC2,
            symbol = ~day,
            mode = "markers",
            text = ~label,
            marker = list(color = "grey", size = 8)) %>% 
  add_trace(type = "scatter",
            x = ~PC1, y = ~PC2,
            color = ~delivery,
            symbol = ~day,
            mode = "markers",
            text = ~label,
            marker = list(size = 10),
            showlegend = F) %>% 
            layout(xaxis = list(title = paste("PC1:", round(pca.val[[1]], 2)*100, "%", sep = " ")),
                   yaxis = list(title = paste("PC2:", round(pca.val[[2]], 2)*100, "%", sep = " ")))

p.1
Sys.setenv("PATH" = paste(Sys.getenv("PATH"), "/anaconda3/bin", sep = .Platform$path.sep)) ## To run orca
orca(p.1, file="../../2-results/PC1_PC2.pdf") # Orca will append to WD (no absolute path)

# if ( !require(RSelenium) ) {
#   install.packages("RSelenium", repos = "https://cloud.r-project.org/")
# }
# p.1 %>% export(file = "test.svg",
#          selenium = RSelenium::rsDriver(browser = "chrome"))
# orca(p.1, file="../../2-results/PC1_PC2.png") # Orca will append to WD (no absolute paths)
```  

```{r pca-2}
p.2 <- plot_ly(data = pca.df) %>%
  add_trace(type = "scatter",
            x = ~PC1, y = ~PC3,
            color = ~delivery, colors = pca.cols,
            symbols = c("1" = "circle", "2" = "cross", "3" = "x", "4" = "square", "6" = "diamond"),
            text = ~label,
            mode = "markers") %>% 
  add_trace(type = "scatter",
            x = ~PC1, y = ~PC3,
            symbol = ~day,
            mode = "markers",
            text = ~label,
            marker = list(color = "grey", size = 8)) %>% 
  add_trace(type = "scatter",
            x = ~PC1, y = ~PC3,
            color = ~delivery,
            symbol = ~day,
            mode = "markers",
            text = ~label,
            marker = list(size = 10),
            showlegend = F) %>% 
            layout(xaxis = list(title = paste("PC1:", round(pca.val[[1]], 2)*100, "%", sep = " ")),
                   yaxis = list(title = paste("PC3:", round(pca.val[[3]], 2)*100, "%", sep = " ")))

p.2
Sys.setenv("PATH" = paste(Sys.getenv("PATH"), "/anaconda3/bin", sep = .Platform$path.sep)) ## To run orca
orca(p.2, file="../../2-results/PC1_PC3.pdf") # Orca will append to WD (no absolute path)

```

```{r pca-3}
p.3 <- plot_ly(data = pca.df) %>%
  add_trace(type = "scatter",
            x = ~PC2, y = ~PC3,
            color = ~delivery, colors = pca.cols,
            symbols = c("1" = "circle", "2" = "cross", "3" = "x", "4" = "square", "6" = "diamond"),
            text = ~label,
            mode = "markers") %>% 
  add_trace(type = "scatter",
            x = ~PC2, y = ~PC3,
            symbol = ~day,
            mode = "markers",
            text = ~label,
            marker = list(color = "grey", size = 8)) %>% 
  add_trace(type = "scatter",
            x = ~PC2, y = ~PC3,
            color = ~delivery,
            symbol = ~day,
            mode = "markers",
            text = ~label,
            marker = list(size = 10),
            showlegend = F) %>% 
            layout(xaxis = list(title = paste("PC2:", round(pca.val[[2]], 2)*100, "%", sep = " ")),
                   yaxis = list(title = paste("PC3:", round(pca.val[[3]], 2)*100, "%", sep = " ")))

# p.3
Sys.setenv("PATH" = paste(Sys.getenv("PATH"), "/anaconda3/bin", sep = .Platform$path.sep)) ## To run orca
orca(p.3, file="../../2-results/PC2_PC3.pdf") # Orca will append to WD (no absolute path)

```

#Differential Gene Expression

##DGE overview

The PCA's may be indicating that, with the exception of day 1 of mosquito transmission, the groups are fairly homogeneous up to day 3. There are small clusters that form, however there seems to be one replicate in each group that bucks the trend in most cases. Doing direct group comparisons means that it's likely to mask siginficant genes each tranmission group uses to fight the infection. If we identify what genes are differentially expressed within each transmission route throughout the time course, and then compare these gene lists, we can see what genes are shared/different between the transmission routes. This will hide some information, such as if a gene/geneset are shared between two transmission routes, but activated at different times, but this approach will narrow down our search in the first instance. 

First of all we need to identify the DGE's in each transmission route, by comparing them to the controls.

```{r dge-res, include = F}
res.list <- list();
design(dds) <- ~ grp;
dds <- DESeq(dds);
resultsNames(dds);  
res.list[["mt.1_vs_naive"]]      <- results(dds, contrast=c("grp","mt.1","naive"));
res.list[["mt.2_vs_naive"]]      <- results(dds, contrast=c("grp","mt.2","naive"));
res.list[["mt.3_vs_naive"]]      <- results(dds, contrast=c("grp","mt.3","naive"));
res.list[["mt.4_vs_naive"]]      <- results(dds, contrast=c("grp","mt.4","naive"));
res.list[["mt.6_vs_naive"]]      <- results(dds, contrast=c("grp","mt.6","naive"));

res.list[["rtmt.1_vs_naive"]]      <- results(dds, contrast=c("grp","rtmt.1","naive"));
res.list[["rtmt.2_vs_naive"]]      <- results(dds, contrast=c("grp","rtmt.2","naive"));
res.list[["rtmt.3_vs_naive"]]      <- results(dds, contrast=c("grp","rtmt.3","naive"));
res.list[["rtmt.4_vs_naive"]]      <- results(dds, contrast=c("grp","rtmt.4","naive"));
res.list[["rtmt.6_vs_naive"]]      <- results(dds, contrast=c("grp","rtmt.6","naive"));

res.list[["sbp.1_vs_naive"]]      <- results(dds, contrast=c("grp","sbp.1","naive"));
res.list[["sbp.2_vs_naive"]]      <- results(dds, contrast=c("grp","sbp.2","naive"));
res.list[["sbp.3_vs_naive"]]      <- results(dds, contrast=c("grp","sbp.3","naive"));
res.list[["sbp.4_vs_naive"]]      <- results(dds, contrast=c("grp","sbp.4","naive"));
res.list[["sbp.6_vs_naive"]]      <- results(dds, contrast=c("grp","sbp.6","naive"));

res.list[['mt.1_vs_sbp.1']]      <- results(dds, contrast=c("grp","mt.1","sbp.1"));
res.list[['mt.2_vs_sbp.2']]      <- results(dds, contrast=c("grp","mt.2","sbp.2"));
res.list[['mt.3_vs_sbp.3']]      <- results(dds, contrast=c("grp","mt.3","sbp.3"));
res.list[['mt.4_vs_sbp.4']]      <- results(dds, contrast=c("grp","mt.4","sbp.4"));
res.list[['mt.6_vs_sbp.6']]      <- results(dds, contrast=c("grp","mt.6","sbp.6"));

res.list[['mt.1_vs_rtmt.1']]      <- results(dds, contrast=c("grp","mt.1","rtmt.1"));
res.list[['mt.2_vs_rtmt.2']]      <- results(dds, contrast=c("grp","mt.2","rtmt.2"));
res.list[['mt.3_vs_rtmt.3']]      <- results(dds, contrast=c("grp","mt.3","rtmt.3"));
res.list[['mt.4_vs_rtmt.4']]      <- results(dds, contrast=c("grp","mt.4","rtmt.4"));
res.list[['mt.6_vs_rtmt.6']]      <- results(dds, contrast=c("grp","mt.6","rtmt.6"));

res.list[['rtmt.1_vs_sbp.1']]      <- results(dds, contrast=c("grp","rtmt.1","sbp.1"));
res.list[['rtmt.2_vs_sbp.2']]      <- results(dds, contrast=c("grp","rtmt.2","sbp.2"));
res.list[['rtmt.3_vs_sbp.3']]      <- results(dds, contrast=c("grp","rtmt.3","sbp.3"));
res.list[['rtmt.4_vs_sbp.4']]      <- results(dds, contrast=c("grp","rtmt.4","sbp.4"));
res.list[['rtmt.6_vs_sbp.6']]      <- results(dds, contrast=c("grp","rtmt.6","sbp.6"));


## Filter results
res.list <- lapply(res.list,function(res){
  res <- as.data.frame(res);
  colnames(res)[colnames(res) %in% "log2FoldChange"] <- "log2FC";
  colnames(res)[colnames(res) %in% "padj"] <- "FDR";
  res$DEG <- res$FDR <= 0.05 &
    abs(res$log2FC)>= 1 & #only filters out logFC ±1
    res$baseMean>=30         &
    !is.na(res$log2FC)       &
    !is.na(res$FDR);
  res;
})
deg.meta <- as.data.frame(cbind(sapply(res.list[1:15],function(x){sum(x$DEG)})))
names(deg.meta) <- "No. DEG"

deg.meta
```

With the comparisons made and genes filtered to identify the DGE's we can see how many genes were identified within in each group.

```{r deg-table}
paged_table(deg.meta, options = list(rows.print = 15))
```

We can then find out the total number of genes used by each transmission group throughout the timecourse.
```{r deg-total}
# Mosquito vs Naive results and DEG
mvn.res <- res.list[grep('^mt.._vs_naive', names(res.list))]
mvn.deg <- lapply(mvn.res, function(df){return(df[df$DEG,])})

# Recently MT vs Naive results and DEG
rtvn.res <- res.list[grep('^rtmt.._vs_naive', names(res.list))]
rtvn.deg <- lapply(rtvn.res, function(df){return(df[df$DEG,])})

# SBP vs Naive results and DEG 
svn.res <- res.list[grep('^sbp.._vs_naive', names(res.list))]
svn.deg <- lapply(svn.res, function(df){return(df[df$DEG,])})

# Genes involved in each transmission route
mt.genes <- unique(unlist(sapply(mvn.deg, row.names)))
rtmt.genes <- unique(unlist(sapply(rtvn.deg, row.names))) 
sbp.genes <- unique(unlist(sapply(svn.deg, row.names))) 
total.genes <- unique(do.call(c, list(mt.genes, rtmt.genes, sbp.genes)))

# Common genes involved in each transmission response
common.genes <- Reduce(intersect, list(mt.genes, rtmt.genes, sbp.genes))


print(paste("MT:", length(mt.genes), "differntially expressed genes identified"))
print(paste("RTMT:", length(rtmt.genes), "differentially expressed genes identified"))
print(paste("SBP:", length(sbp.genes), "differentially expressed genes identified"))
print(paste("Total number of differentially expressed genes identified across all groups: ", length(total.genes), "genes"))

```


##MT Day 1 -3
### Gene List

The general DGE analysis reveals that the MT group is transcriptionally different from Naive from Day 1-3, where other groups only show transcriptional differences from Day 4 onwards. The question remains is to what is happening in this transmission group that make
```{r}
mvn.early <- mvn.deg[1:3]
mvn.early.list <- unique(unlist(sapply(mvn.early, row.names)))

## Comment out during markdown rendering
# write.table(x = mvn.early.list, file = paste(results.dir, "mvn_early_deg.txt", sep=""),
#             sep = "\t",
#             row.names = F,
#             quote = F)


# Panther DF
panther.df <- read.table(file = paste(results.dir, "mvn_pantherGeneList.txt", sep=""),
                         sep = "\t",
                         quote = "",
                         header = F,
                         col.names = c("panther_id", "gene_id", "gene_name", "panther_family", "panther_protein", "species")) %>% 
  separate_rows(gene_id, sep = ",", convert = TRUE) %>% 
  separate_rows(panther_protein, sep = ";")
  
panther.mvn <- panther.df[panther.df$gene_id %in% mvn.early.list, ]

```




```{r innate-db}
innate.db <- read.table(file = paste(results.dir, "mvn_InnateDB_genes.txt", sep=""),
                         sep = "\t",
                         quote = "",
                         header = T)

unique(innate.db$Pathway.Name)
innate.path.df <- as.data.frame(table(innate.db$Pathway.Name), col.names = c("Pathway", "Freq")) 
innate.path.df[order(innate.path.df$Freq, decreasing = T), ]

mvn.innate <- unique(innate.db$Query.Xref)
mvn.innate.ensembl <- gtf.dat$gene_id[gtf.dat$gene_name %in% mvn.innate]
tpm.db["ENSMUSG00000021264", ]
```
 
```{r innate-heatmap}
heat.dat   <- as.matrix(assay(vsd)[mvn.innate.ensembl, ]);
heat.dat   <- t(apply(heat.dat,1,function(x){((x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE))}));
rownames(heat.dat) <- gtf.dat[mvn.innate.ensembl, "gene_name"];

heat.dat <- heat.dat[ , grep("^mt|naive", colnames(heat.dat))]

#order so days are together 
heat.list <- colnames(heat.dat);
heat.list <- as.data.frame(heat.list);
heat.list$day <- ifelse(grepl("naive", heat.list$heat.list),
                        0,
                        str_match(string=heat.list$heat.list, pattern= "\\.d(.+)\\.")[,2]);
heat.list$delivery <- ifelse(grepl("naive", heat.list$heat.list),
                             "naive",
                             str_match(string=heat.list$heat.list, pattern="^(.*)\\..*\\.")[,2]);
heat.list$delivery <- ordered(factor(heat.list$delivery, levels= c('naive', 'sbp', 'rtmt', 'mt')))
heat.list$rep <- as.integer(str_match(string=heat.list$heat.list, pattern=".*r(.)$")[,2]);
heat.list <- heat.list[order(as.numeric(heat.list$day), as.numeric(as.factor(heat.list$delivery)), heat.list$rep),];
heat.list
heat.list <- as.vector(heat.list$heat.list);


heat.dat <- heat.dat[,heat.list] # Use the same col order as the for the zscore heatmap


ht <- Heatmap(heat.dat,
              name = "z-score",
              show_row_names          = TRUE,
              show_column_names       = TRUE,
              column_names_side       = "bottom",
              show_column_dend        = FALSE,
              cluster_columns         = FALSE,
              rect_gp                 = gpar(col = "white", lwd = 1),
              row_dend_width          = unit(30, "mm"),
              column_dend_height      = unit(30, "mm"),
              show_heatmap_legend     = TRUE,
              clustering_distance_rows = 'pearson',
              column_names_max_height = unit(8, "cm"),
              row_names_gp            = gpar(fontsize = 6),
              column_names_gp         = gpar(fontsize = 10),
              # split                   = split.vec, #This would use vector I created above to split and cluster the genes
              # top_annotation          = ha1,
              # bottom_annotation       = anno_bottom,
              col                     = colorRamp2(c(min(heat.dat)-0.5, 0, max(heat.dat)+0.5), rev(colorRampPalette(brewer.pal(9, "RdBu"))(3))));


pdf(paste(results.dir,"mvn_innate_hm.pdf",sep=''),width=15,height=30);
ht
decorate_heatmap_body('z-score', c(for(d in seq(0, length(heat.list), 4)){
  grid.lines(x=c(d/24, d/24), y=c(0,1), gp= gpar(col = 'black', lty="solid", lwd=2))},
  grid.lines(y=c(0,0), x=c(0,1), gp= gpar(col = 'black', lty='solid', lwd=2)),
  grid.lines(y=c(1,1), x=c(0,1), gp= gpar(col = 'black', lty='solid', lwd=2))), slice = 1)
dev.off();
```

```{r}
mvn.1.3 <- lapply(mvn.early, row.names)

mvn.1.3.innate <- lapply(mvn.1.3, function(x){
  x <- intersect(x, mvn.innate.ensembl) 
  x
})

mvn.1.3.innate[[1]]
```

```{r}
heat.dat   <- as.matrix(tpm.db[mvn.1.3.innate[[1]], ]);
heat.dat   <- log2(heat.dat + 1)
# heat.dat   <- t(apply(heat.dat,1,function(x){((x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE))}));
rownames(heat.dat) <- gtf.dat[mvn.1.3.innate[[1]], "gene_name"];

heat.dat <- heat.dat[ , grep("^mt|naive", colnames(heat.dat))]

#order so days are together 
heat.list <- colnames(heat.dat);
heat.list <- as.data.frame(heat.list);
heat.list$day <- ifelse(grepl("naive", heat.list$heat.list),
                        0,
                        str_match(string=heat.list$heat.list, pattern= "\\.d(.+)\\.")[,2]);
heat.list$delivery <- ifelse(grepl("naive", heat.list$heat.list),
                             "naive",
                             str_match(string=heat.list$heat.list, pattern="^(.*)\\..*\\.")[,2]);
heat.list$delivery <- ordered(factor(heat.list$delivery, levels= c('naive', 'sbp', 'rtmt', 'mt')))
heat.list$rep <- as.integer(str_match(string=heat.list$heat.list, pattern=".*r(.)$")[,2]);
heat.list <- heat.list[order(as.numeric(heat.list$day), as.numeric(as.factor(heat.list$delivery)), heat.list$rep),];
heat.list
heat.list <- as.vector(heat.list$heat.list);


heat.dat <- heat.dat[,heat.list] # Use the same col order as the for the zscore heatmap


ht <- Heatmap(heat.dat,
              name = "log(tpm+1)",
              show_row_names          = TRUE,
              show_column_names       = TRUE,
              column_names_side       = "bottom",
              show_column_dend        = FALSE,
              cluster_columns         = FALSE,
              rect_gp                 = gpar(col = "white", lwd = 1),
              row_dend_width          = unit(30, "mm"),
              column_dend_height      = unit(30, "mm"),
              show_heatmap_legend     = TRUE,
              clustering_distance_rows = 'pearson',
              column_names_max_height = unit(8, "cm"),
              row_names_gp            = gpar(fontsize = 6),
              column_names_gp         = gpar(fontsize = 10),
              # split                   = split.vec, #This would use vector I created above to split and cluster the genes
              # top_annotation          = ha1,
              # bottom_annotation       = anno_bottom,
              col                     = viridis::viridis(100));


pdf(paste(results.dir,"mvn_innate_d1_tpm.pdf",sep=''),width=15,height=30);
ht
decorate_heatmap_body('log(tpm+1)', c(for(d in seq(0, length(heat.list), 4)){
  grid.lines(x=c(d/24, d/24), y=c(0,1), gp= gpar(col = 'black', lty="solid", lwd=2))},
  grid.lines(y=c(0,0), x=c(0,1), gp= gpar(col = 'black', lty='solid', lwd=2)),
  grid.lines(y=c(1,1), x=c(0,1), gp= gpar(col = 'black', lty='solid', lwd=2))), slice = 1)
dev.off();

```
```{r}
ensembl <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")
attributes <- listAttributes(ensembl)
attributes
#ensembl
# mvn.early.biomart <-  getBM(attributes = c("ensembl_gene_id","family", "family_description", "pfam", "interpro", "interpro_short_description"),
mvn.early.fam <-  getBM(attributes = c("ensembl_gene_id","family", "family_description"),
      filters = "ensembl_gene_id",
      values = mvn.early.list,
      mart = ensembl) %>% filter(family != "")

mvn.fam.df <- as.data.frame(table(mvn.early.fam$family))
mvn.fam.df <- mvn.fam.df[order(mvn.fam.df$Freq, decreasing = T), ]
mvn.fam.df$description <- sapply(mvn.fam.df$Var1, function(x){
  df <- unique(mvn.early.fam[mvn.early.fam$family %in% x, "family_description"])
  df
})
mvn.fam.df$Var1 <- factor(mvn.fam.df$Var1, levels = mvn.fam.df$Var1)
View(mvn.fam.df)

#pfam
mvn.early.pfam <-  getBM(attributes = c("ensembl_gene_id","pfam"),
      filters = "ensembl_gene_id",
      values = mvn.early.list,
      mart = ensembl) %>% filter(pfam != "")
mvn.pfam.df <- as.data.frame(table(mvn.early.pfam$pfam))
mvn.pfam.df <- mvn.pfam.df[order(mvn.pfam.df$Freq, decreasing = T), ]


#interpro
mvn.early.interpro <-  getBM(attributes = c("ensembl_gene_id","interpro", "interpro_description"),
      filters = "ensembl_gene_id",
      values = mvn.early.list,
      mart = ensembl) %>% filter(interpro != "")

mvn.interpro.df <- as.data.frame(table(mvn.early.interpro$interpro))
mvn.interpro.df <- mvn.interpro.df[order(mvn.interpro.df$Freq, decreasing = T), ]
mvn.interpro.df$desc <- sapply(mvn.interpro.df$Var1, function(x){
  df <- unique(mvn.early.interpro[mvn.early.interpro$interpro %in% x, "interpro_description"])
  df
})
mvn.interpro.df$Var1 <- factor(mvn.interpro.df$Var1,levels = mvn.interpro.df$Var1)
View(mvn.early.interpro)
View(mvn.interpro.df)
View(mvn.pfam.df)
View(mvn.fam.df)
mvn.early.interpro$interpro <- factor(mvn.early.interpro$interpro, levels = mvn.interpro.df$Var1)
mvn.early.fam$family <- factor(mvn.early.fam$family, levels = mvn.fam.df$Var1)

```
 

```{r}
tf.db <- read.table("/Users/alderc/1-projects/9-Data/tf_db_riken.txt",
                    sep = "\t",
                    header = T,
                    quote = "")

rownames(tf.db) <- tf.db$Symbol

str_to_upper(mvn.innate)
mvn.tf <- tf.db[mvn.innate, ] %>% drop_na()


tf.trrust.db <- read.table("/Users/alderc/1-projects/9-Data/trrust_rawdata.mouse.tsv",
                           sep = "\t",
                           header = T,
                           quote = "")
mvn.trrust.df <- tf.trrust.db[tf.trrust.db$TF %in% mvn.innate, ]

View(mvn.trrust.tf)

mvn.trrust.tf <- unique(mvn.trrust.tf$TF)

mvn.early.names <- gtf.dat[mvn.early.list, "gene_name"]
mvn.early.tf.targets <- mvn.trrust.df[mvn.trrust.df$Target %in% mvn.early.names, ]
unique(mvn.early.tf.targets$Target)

```

```{r}
mvn.df <- data.frame(gene_id = mvn.early.list,
                     gene_symbol = mvn.early.names,
                     row.names = 1)

mvn.df$Day_1 <- ifelse(rownames(mvn.df) %in% rownames(mvn.early[[1]]), "Yes", "No")
mvn.df$Day_2 <- ifelse(rownames(mvn.df) %in% rownames(mvn.early[[2]]), "Yes", "No")
mvn.df$Day_3 <- ifelse(rownames(mvn.df) %in% rownames(mvn.early[[3]]), "Yes", "No")
mvn.df$innate <- ifelse(rownames(mvn.df) %in% mvn.innate.ensembl, "Yes", "No")
mvn.df$tf <- ifelse(mvn.df$gene_symbol %in% mvn.trrust.tf, "Yes", "No")
mvn.df$inter <- sapply(rownames(mvn.df), function(x){
  df <- mvn.early.interpro[mvn.early.interpro$ensembl_gene_id %in% x, ]
  df <- df[order(df$interpro), "interpro_description"]
  df[1]
})
mvn.df$ensembl_fam <- sapply(rownames(mvn.df), function(x){
  df <- mvn.early.fam[mvn.early.fam$ensembl_gene_id %in% x, ]
  df <- df[order(df$family), "family_description"]
  df[1]
})
mvn.df$rtmt <- sapply(rownames(mvn.df), function(g){
for (i in c(1,2,3,4,6)){
  tmp <- NULL
  x <- paste("rtmt.", i, "_vs_naive", sep = "")
  df <- rvn.deg[[x]]
  df <- df[g, ] %>% drop_na()
  if(dim(df)[1] == 1){
    tmp <- i
    break
  }
}
  if(is.null(tmp)){
    tmp <- NA
  }
  return(tmp)
})
mvn.df$sbp <- sapply(rownames(mvn.df), function(g){
for (i in c(1,2,3,4,6)){
  tmp <- NULL
  x <- paste("sbp.", i, "_vs_naive", sep = "")
  df <- svn.deg[[x]]
  df <- df[g, ] %>% drop_na()
  if(dim(df)[1] == 1){
    tmp <- i
    break
  }
}
  if(is.null(tmp)){
    tmp <- NA
  }
  return(tmp)
})

View(mvn.df)

rvn.deg <- deg.list[6:10]
svn.deg <- deg.list[11:15]
m.genes <- rownames(mvn.df)
m.genes[1]
sapply(c(m.genes[1:5], "ENSMUSG00000064844"), function(g){
for (i in c(1,2,3,4,6)){
  tmp <- NULL
  x <- paste("rtmt.", i, "_vs_naive", sep = "")
  df <- rvn.deg[[x]]
  df <- df[g, ] %>% drop_na()
  if(dim(df)[1] == 1){
    tmp <- i
    break
  }
}
  if(is.null(tmp)){
    tmp <- NA
  }
  return(tmp)
})

```
```{r innate-heatmap}
  # heat.dat   <- as.matrix(assay(vsd)[mvn.early.list, ]);
# sample.remove <- grep("d4|d6", colnames(heat.dat), value = TRUE)
# heat.dat <- heat.dat[ , !colnames(heat.dat) %in% sample.remove]
# heat.dat   <- t(apply(heat.dat,1,function(x){((x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE))}));

heat.dat   <- as.matrix(tpm.db[mvn.early.list, ]);
heat.dat   <- log(heat.dat + 1)
sample.remove <- grep("d4|d6", colnames(heat.dat), value = TRUE)
heat.dat <- heat.dat[ , !colnames(heat.dat) %in% sample.remove]
rownames(heat.dat) <- gtf.dat[rownames(heat.dat), "gene_name"];

heat.dat



#order so days are together 
heat.list <- colnames(heat.dat);
heat.list <- as.data.frame(heat.list);
heat.list$day <- ifelse(grepl("naive", heat.list$heat.list),
                        0,
                        str_match(string=heat.list$heat.list, pattern= "\\.d(.+)\\.")[,2]);
heat.list$delivery <- ifelse(grepl("naive", heat.list$heat.list),
                             "naive",
                             str_match(string=heat.list$heat.list, pattern="^(.*)\\..*\\.")[,2]);
heat.list$delivery <- ordered(factor(heat.list$delivery, levels= c('naive', 'mt', 'sbp', 'rtmt')))
heat.list$rep <- as.integer(str_match(string=heat.list$heat.list, pattern=".*r(.)$")[,2]);
heat.list <- heat.list[order(as.numeric(heat.list$day), as.numeric(as.factor(heat.list$delivery)), heat.list$rep),];
heat.list
heat.list <- as.vector(heat.list$heat.list);


heat.dat <- heat.dat[,heat.list] # Use the same col order as the for the zscore heatmap


ht <- Heatmap(heat.dat,
              name = "log(tpm+1)",
              show_row_names          = TRUE,
              show_column_names       = TRUE,
              column_names_side       = "bottom",
              show_column_dend        = FALSE,
              show_row_dend           = FALSE, 
              cluster_columns         = FALSE,
              rect_gp                 = gpar(col = "white", lwd = 1),
              row_dend_width          = unit(30, "mm"),
              column_dend_height      = unit(30, "mm"),
              show_heatmap_legend     = TRUE,
              clustering_distance_rows = 'pearson',
              column_names_max_height = unit(8, "cm"),
              row_names_gp            = gpar(fontsize = 4),
              column_names_gp         = gpar(fontsize = 10),
              # split                   = split.vec, #This would use vector I created above to split and cluster the genes
              # top_annotation          = ha1,
              # bottom_annotation       = anno_bottom,
              # col                     = colorRamp2(c(min(heat.dat), 0, max(heat.dat)), rev(colorRampPalette(brewer.pal(9, "RdBu"))(3))));
              col                     = viridis::viridis(100));

group <- c(0,4,16,28,40)
pdf(paste(results.dir,"mvn_early_hm_w_naive_tpm_viridis.pdf",sep=''),width=15,height=30);
ht
decorate_heatmap_body("log(tpm+1)", c(for(d in seq(0, length(heat.list), 4)){
  grid.lines(x=c(d/40, d/40), y=c(0,1), gp= gpar(col = 'black', lty="solid", lwd=2))},
  grid.lines(y=c(0,0), x=c(0,1), gp= gpar(col = 'black', lty='solid', lwd=2)),
  grid.lines(y=c(1,1), x=c(0,1), gp= gpar(col = 'black', lty='solid', lwd=2))), slice = 1)
dev.off();

heat.dat
```

